Explainability and Adversarial Robustness for RNNs

### 2020-07-08 paper/presentation/slides.tex
Changed slides

\section{Intrusion Detection using RNNs}
\begin{frame}{Intrusion Detection using RNNs}
\begin{itemize}
\item Detect attacks before they are over
\item Classifier obtains all available information
\end{itemize}
\vspace{10pt}
\centering
\includegraphics[width=0.7\columnwidth]{figures/rnn.pdf}
\end{frame}

@@

\begin{frame}{Finding Adversarial Samples}
\begin{itemize}
\item Carlini-Wagner {\footnotesize ($X$...orig. sample, $Z$...logit output, $\kappa,\delta$...parameters) \hspace*{-1cm} }
\begin{equation*} \label{eq:carliniWagner}
\argmin_{\tilde X}d(X,\tilde X) + \kappa  \max(Z(\tilde X), \delta)
\end{equation*}
\item FGSM {\footnotesize ($L$...neural network loss, $\epsilon$...parameter) }
\begin{equation*}
\tilde X = X + \epsilon \, \text{sgn}( \nabla_X L(X))
\end{equation*}
\item $L_\infty$-bounded Projected Gradient Descent

@@

\begin{frame}{The Adversarial Robustness Score}
\begin{itemize}
\item Quantify ease of generating adversarial samples
\item Carlini-Wagner {\footnotesize ($X$...orig. sample, $Z$...logit output, $\kappa,\delta$...parameters) \hspace*{-1cm} }
\begin{equation*} \label{eq:carliniWagner}
\argmin_{\tilde X} d(X,\tilde X) + \kappa  \max(Z(\tilde X), \delta)
\end{equation*}
\item ARS $\approx$ avg. distance of the 50\% successful adversarial samples with lowest distance
\end{itemize}
\end{frame}

@@

\section{Defenses}
\begin{frame}{Defenses}
\begin{itemize}
\item Omit manipulable features
\begin{itemize}
\item Either all packet lengths and IATs
\item or just in one direction
\end{itemize}
$\rightarrow$ Surprisingly good accuracy \vspace{10pt}
\item Adversarial Training

@@

\section{Conclusions}
\begin{frame}{Conclusions}
\begin{itemize}
\item RNNs usable for IDSs
\item Adversarial ML algorithms work
\begin{itemize}
\item Defenses work, too
\item ARS for quantifying robustness
\end{itemize}
\item Explainability
\begin{itemize}
\item Feature importance and sensitivity, dependent on purpose 
\item Visualization, e.g. sequential PDPs
\item Find packets which are characteristic for attack
\end{itemize}
\end{itemize}
\end{frame}
% -------------
% Last page
% -------------
\makelastslide

>>>>

\section{An RNN for Intrusion Detection}
\begin{frame}{Using a Recurrent Neural Network}
\begin{itemize}
\item Detect attacks before they are over
\item Classifier obtains all available information
\end{itemize}
\vspace{10pt}
\centering
\includegraphics[width=0.8\columnwidth]{figures/rnn.pdf}
\end{frame}

@@

\begin{frame}{Finding Adversarial Samples}
\begin{itemize}
\item Carlini-Wagner {\footnotesize ($X$...orig.~sample; $Z$...logit output; $\kappa,\delta$...parameters) \hspace*{-1cm} }
\begin{equation*} \label{eq:carliniWagner}
\argmin_{\tilde X}d(X,\tilde X) + \kappa  \max(Z(\tilde X), \delta)
\end{equation*}
\item FGSM {\footnotesize ($L$...neural network loss; $\epsilon$...parameter) }
\begin{equation*}
\tilde X = X + \epsilon \, \text{sgn}( \nabla_X L(X))
\end{equation*}
\item $L_\infty$-bounded Projected Gradient Descent

@@

\begin{frame}{The Adversarial Robustness Score}
\begin{itemize}
\item Quantify ease of generating adversarial samples
\item Carlini-Wagner {\footnotesize ($X$...orig.~sample; $Z$...logit output; $\kappa,\delta$...parameters) \hspace*{-1cm} }
\begin{equation*} \label{eq:carliniWagner}
\argmin_{\tilde X} d(X,\tilde X) + \kappa  \max(Z(\tilde X), \delta)
\end{equation*}
\item ARS $\approx$ avg. distance of the 50\% successful adversarial samples with lowest distance
\end{itemize}
\end{frame}

@@

\section{Defenses}
\begin{frame}{Defenses}
\begin{itemize}
\item Omit manipulable features
\begin{itemize}
\item Either omit all packet lengths and IATs
\item or retain features in non-manipulable direction
\end{itemize}
$\rightarrow$ Surprisingly good accuracy \vspace{10pt}
\item Adversarial Training

@@

\section{Conclusions}
\begin{frame}{Conclusions}
\begin{itemize}
\item RNNs usable for IDSs
\item Adversarial ML algorithms work
\begin{itemize}
\item Defenses work, too
\item \textit{Adversarial Robustness Score} for quantifying robustness
\end{itemize}
\item Explainability
\begin{itemize}
\item Feature importance and sensitivity, dependent on purpose 
\item Visualization\\$\rightarrow$ e.g. \textit{sequential Partial Dependence Plots}
\item Find packets which are characteristic for attack
\end{itemize}
\end{itemize}
\end{frame}
% -------------
% Last page
% -------------
\makelastslide
